---
title: "Thesis_Analysis"
author: "Jade Benson"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lme4)
require(broom.mixed)
library(MLmetrics)

#rolling origin 
#https://www.openforecast.org/adam/rollingOrigin.html
#https://rdrr.io/cran/greybox/man/ro.html
#library(greybox)
#this only works with explicitly forecasting models 

```
In this notebook, I run Poisson regressions using the weather, sociodemographics, and monthly fall rates from US counties from 2009 - 2015. I am basing this on this analysis: 
https://github.com/lin-lab/COVID-Health-Disparities/blob/main/Scripts/Case%20Rate%20Total%20Models.R
In this they add state as a variable which is just the state name. 
They also add CountyI which is just a unique row indicator for each county: 1001 is 1, 1002 is 2, and 1005 is 3 (since it's the next one) but the first Alaska one is 68. 

```{r load-data}
no_missing_regions_df <- read_csv("/Users/jadebenson/Documents/Thesis/no_missing_data.csv")

head(no_missing_regions_df)

#add in column with unique number for each county
fipscounty <- no_missing_regions_df['fipscounty']

unique_counties <- fipscounty %>%
  distinct(fipscounty)

unique_counties$County_ID <- seq.int(nrow(unique_counties))

#merge this back in 
data <- merge(no_missing_regions_df, unique_counties, by.x = 'fipscounty', by.y = 'fipscounty')

head(data)
#worked!

length(data$scaled_fall_rate)

#maybe can just cast the fall rates as integers? 
data$int_fall_rate = as.integer(data$scaled_fall_rate)

#split into training and testing sets 
typeof(data$date)

data <- data %>%
  mutate(year = substr(date, 1, 4) )

data <- data %>%
  mutate(test_set = ifelse((year == '2014') | (year == '2015'), 1, 0))
  
train_data <-  data[ which(data$test_set==0), ]
length(train_data$month_count)

test_data <- data[ which(data$test_set==1), ]
length(test_data$month_count)

geographic_vars <- c('pop_65_up', 'STATE', 'County_ID')
weather_vars <- c('tavg_mean', 'tmin_min', 'tmax_max', 'prcp_max', 'snow_max', 'wspd_max')
demo_vars <- c('prop_65_85_up', 'prop_female', 'prop_white_nh', 'pop_density')
econ_vars <- c('median_income', 'gini_index', 'prop_poor')
full_vars <- c(weather_vars, demo_vars, econ_vars)

```


Functions 

```{r weather-fall-rates-func}

weather_function <- function(train_df, test_df, weather_vector){
  
  weather_train <- train_df %>% 
    select(c('int_fall_rate', geographic_vars,  weather_vector)) %>% 
    mutate_at(weather_vector, funs(c(scale(.))))
  
  weather_test <- test_df %>% 
    select(c('int_fall_rate', geographic_vars,  weather_vector)) %>% 
    mutate_at(weather_vector, funs(c(scale(.))))
  
  #Model Fit
  weather_rate_model <- glmer(formula = int_fall_rate ~ offset(log(pop_65_up)) + STATE +  (1 | County_ID) +
                  tavg_mean + tmin_min + tmax_max +  
                  prcp_max + snow_max + wspd_max, 
                family=poisson(),
                nAGQ=0,
                data = weather_train)
  
  
  summary(weather_rate_model)
  confint = tidy(weather_rate_model,conf.int=TRUE,exponentiate=TRUE,effects="fixed")
  multivariable = confint[, c(2, 3, 6, 7, 8)] %>% print(n = Inf)
  
  #train evaluation
  weather_train_predictions= predict(weather_rate_model, select(weather_train, -c('int_fall_rate')) )
  weather_train_MAPE = MAPE(weather_train$int_fall_rate, weather_train_predictions)
  print(paste0("Train MAPE: ", weather_train_MAPE))
  
  
  #test evaluation
   weather_test_predictions=predict(weather_rate_model, select(weather_test, -c('int_fall_rate')), allow.new.levels = TRUE)
  weather_test_MAPE = MAPE(weather_test$int_fall_rate, weather_test_predictions)
  print(paste0("Testing MAPE: ", weather_test_MAPE))
}


```



```{r demo-model-func}

demo_function <- function(train_df, test_df, demo_vector){
  
  demo_train <- train_df %>% 
    select(c('int_fall_rate', geographic_vars,  demo_vector)) %>% 
    mutate_at(demo_vector, funs(c(scale(.))))
  
  demo_test <- test_df %>% 
    select(c('int_fall_rate', geographic_vars,  demo_vector)) %>% 
    mutate_at(demo_vector, funs(c(scale(.))))
  
  #Model Fit
  demo_rate_model <- glmer(formula = int_fall_rate ~ offset(log(pop_65_up)) + STATE +  (1 | County_ID) +
                  prop_65_85_up + prop_female + prop_white_nh + pop_density, 
                family=poisson(),
                nAGQ=0,
                data = demo_train)
  
  
  summary(demo_rate_model)
  confint = tidy(demo_rate_model,conf.int=TRUE,exponentiate=TRUE,effects="fixed")
  multivariable = confint[ , c(2, 3, 6, 7, 8)] %>% print(n = Inf)
  
  #train evaluation
  demo_train_predictions= predict(demo_rate_model, select(demo_train, -c('int_fall_rate')) )
  demo_train_MAPE = MAPE(demo_train$int_fall_rate, demo_train_predictions)
  print(paste0("Train MAPE: ", demo_train_MAPE))
  
  
  #test evaluation
   demo_test_predictions=predict(demo_rate_model, select(demo_test, -c('int_fall_rate')), allow.new.levels = TRUE)
  demo_test_MAPE = MAPE(demo_test$int_fall_rate, demo_test_predictions)
  print(paste0("Testing MAPE: ", demo_test_MAPE))
}

```


```{r econ-fall-rates}

econ_function <- function(train_df, test_df, econ_vector){
  
  econ_train <- train_df %>% 
    select(c('int_fall_rate', geographic_vars,  econ_vector)) %>% 
    mutate_at(econ_vector, funs(c(scale(.))))
  
  econ_test <- test_df %>% 
    select(c('int_fall_rate', geographic_vars,  econ_vector)) %>% 
    mutate_at(econ_vector, funs(c(scale(.))))
  
  #Model Fit
  econ_rate_model <- glmer(formula = int_fall_rate ~ offset(log(pop_65_up)) + STATE +  (1 | County_ID) +
                  median_income + gini_index + prop_poor , 
                family=poisson(),
                nAGQ=0,
                data = econ_train)
  
  
  summary(econ_rate_model)
  confint = tidy(econ_rate_model,conf.int=TRUE,exponentiate=TRUE,effects="fixed")
  sum_len = length(summary(econ_rate_model))
  var_len = length(econ_vars)
  multivariable = confint[ ,c(2, 3, 6, 7, 8)] %>% print(n = Inf)
  
  
  #train evaluation
  econ_train_predictions= predict(econ_rate_model, select(econ_train, -c('int_fall_rate')) )
  econ_train_MAPE = MAPE(econ_train$int_fall_rate, econ_train_predictions)
  print(paste0("Train MAPE: ", econ_train_MAPE))
  
  
  #test evaluation
   econ_test_predictions=predict(econ_rate_model, select(econ_test, -c('int_fall_rate')), allow.new.levels = TRUE)
  econ_test_MAPE = MAPE(econ_test$int_fall_rate, econ_test_predictions)
  print(paste0("Testing MAPE: ", econ_test_MAPE))
}

```


```{r full-func}

full_function <- function(train_df, test_df, full_vector){
  
  full_train <- train_df %>% 
    select(c('int_fall_rate', geographic_vars,  full_vector)) %>% 
    mutate_at(full_vector, funs(c(scale(.))))
  
  full_test <- test_df %>% 
    select(c('int_fall_rate', geographic_vars,  full_vector)) %>% 
    mutate_at(full_vector, funs(c(scale(.))))
  
  #Model Fit
  full_rate_model <- glmer(formula = int_fall_rate ~ offset(log(pop_65_up)) + STATE +  (1 | County_ID) +
                  tavg_mean + tmin_min + tmax_max + prcp_max + snow_max + wspd_max + 
                  prop_65_85_up + prop_female + prop_white_nh + pop_density + 
                  median_income + gini_index + prop_poor , 
                family=poisson(),
                nAGQ=0,
                data = full_train)
  
  
  summary(full_rate_model)
  confint = tidy(full_rate_model,conf.int=TRUE,exponentiate=TRUE,effects="fixed")
  multivariable = confint[ ,c(2, 3, 6, 7, 8)] %>% print(n = Inf)
  
  
  #train evaluation
  full_train_predictions= predict(full_rate_model, select(full_train, -c('int_fall_rate')) )
  full_train_MAPE = MAPE(full_train$int_fall_rate, full_train_predictions)
  print(paste0("Train MAPE: ", full_train_MAPE))
  
  
  #test evaluation
   full_test_predictions=predict(full_rate_model, select(full_test, -c('int_fall_rate')), allow.new.levels = TRUE)
  full_test_MAPE = MAPE(full_test$int_fall_rate, full_test_predictions)
  print(paste0("Testing MAPE: ", full_test_MAPE))
}

```



National Models
```{r national-models}
weather_function(train_df = train_data, test_df = test_data, weather_vector = weather_vars)
demo_function(train_df = train_data, test_df = test_data, demo_vector = demo_vars)
econ_function(train_df = train_data, test_df = test_data, econ_vector = econ_vars)
full_function(train_df = train_data, test_df = test_data, full_vector = full_vars)

```


Regional Models 

```{r regional-dfs}

west_train <- train_data[which(train_data$region == 'west'), ]
west_test <- test_data[which(test_data$region == 'west'), ]
#good same length as in Python
length(west_train$int_fall_rate) + length(west_test$int_fall_rate)


midwest_train <- train_data[which(train_data$region == 'midwest'), ]
midwest_test <- test_data[which(test_data$region == 'midwest'), ]
#good 
length(midwest_train$int_fall_rate) + length(midwest_test$int_fall_rate)


south_train <- train_data[which(train_data$region == 'south'), ]
south_test <- test_data[which(test_data$region == 'south'), ]
#good 
length(south_train$int_fall_rate) + length(south_test$int_fall_rate)


northeast_train <- train_data[which(train_data$region == 'northeast'), ]
northeast_test <- test_data[which(test_data$region == 'northeast'), ]
#good 
length(northeast_train$int_fall_rate) + length(northeast_test$int_fall_rate)

```

```{r regional-models}

#weather
weather_function(train_df = west_train, test_df = west_test, weather_vector = weather_vars)
weather_function(train_df = midwest_train, test_df = midwest_test, weather_vector = weather_vars)
weather_function(train_df = south_train, test_df = south_test, weather_vector = weather_vars)
weather_function(train_df = northeast_train, test_df = northeast_test, weather_vector = weather_vars)


#demographics
demo_function(train_df = west_train, test_df = west_test, demo_vector = demo_vars)
demo_function(train_df = midwest_train, test_df = midwest_test, demo_vector = demo_vars)
demo_function(train_df = south_train, test_df = south_test, demo_vector = demo_vars)
demo_function(train_df = northeast_train, test_df = northeast_test, demo_vector = demo_vars)


#socioeconomics
econ_function(train_df = west_train, test_df = west_test, econ_vector = econ_vars)
econ_function(train_df = midwest_train, test_df = midwest_test, econ_vector = econ_vars)
econ_function(train_df = south_train, test_df = south_test, econ_vector = econ_vars)
econ_function(train_df = northeast_train, test_df = northeast_test, econ_vector = econ_vars)


#full
full_function(train_df = west_train, test_df = west_test, full_vector = full_vars)
full_function(train_df = midwest_train, test_df = midwest_test, full_vector = full_vars)
full_function(train_df = south_train, test_df = south_test, full_vector = full_vars)
full_function(train_df = northeast_train, test_df = northeast_test, full_vector = full_vars)


```
